---
title: "An introduction to biodbChebi"
author: "Firstname of author Lastname of author"
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('biodbChebi')`"
vignette: |
  %\VignetteIndexEntry{Introduction to the biodbChebi package.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: false
  BiocStyle::pdf_document: default
---

# Purpose

biodbChebi is a *biodb* extension package that implements a connector to
Foo database.

# Initialization

The first step in using *biodbChebi*, is to create an instance of the biodb
class `BiodbMain` from the main *biodb* package. This is done by calling the
constructor of the class:
```{r, results='hide'}
mybiodb <- biodb::newInst()
```
During this step the configuration is set up, the cache system is initialized
and extension packages are loaded.

We will see at the end of this vignette that the *biodb* instance needs to be
terminated with a call to the `terminate()` method.

# Creating a connector to Foo database

In *biodb* the connection to a database is handled by a connector instance that
you can get from the factory.
biodbChebi implements a connector to a local database, thus when creating an
instance you must provide a URL that points to your database:
```{r}
chebi <- mybiodb$getFactory()$createConn('chebi')
```

# Requesting entries

Using accession numbers, we request entries from ChEBI:
```{r}
entries <- chebi$getEntry(c('2528', '17053', '15440'))
```

# Get entry fields

Getting the values of entry fields are done using `getFieldValue()` method.
Here we retrieve the SMILES fields for all ChEBI entries obtained previously:
```{r}
vapply(entries, function(e) e$getFieldValue('smiles'), FUN.VALUE = '')
```

# Get data frame

We can convert an entry into a data frame:
```{r}
entries[[1]]$getFieldsAsDataframe()
```

Building a data frame for a list of entries is also possible:
```{r}
mybiodb$entriesToDataframe(entries, fields = c('accession', 'formula', 'molecular.mass', 'inchikey', 'kegg.compound.id'))
```

# Search by name and mass

Searching by name and/or mass is done through the `searchCompound()`.

Searching by name:
```{r}
chebi$searchCompound(name = 'aspartic', max.results = 3)
```

Searching by mass:
```{r}
chebi$searchCompound(mass = 133, mass.field = 'molecular.mass', mass.tol = 0.3, max.results = 3)
```

Searching by name and mass:
```{r}
ids <- chebi$searchCompound(name = 'aspartic', mass = 133, mass.field = 'molecular.mass', mass.tol = 0.3, max.results = 3)
```
Display results in data frame:
```{r}
mybiodb$entriesToDataframe(chebi$getEntry(ids), fields = c('accession', 'molecular.mass', 'name'))
```

# Convert CAS IDs

Converting CAS IDs to ChEBI IDs:
```{r}
chebi$convCasToChebi(c('87605-72-9', '51-41-2'))
```

If more than one ChEBI ID is found for a CAS ID, then a list of character vectors is returned:
```{r}
chebi$convCasToChebi('14215-68-0')
```
This behaviour can be made the default one by setting `simplify` parameter to `FALSE`.

# Convert InChI

The method is similar to `convCasToChebi()`.

Converting InChI to ChEBI IDs:
```{r}
chebi$convInchiToChebi('InChI=1S/C8H11NO3/c9-4-8(12)5-1-2-6(10)7(11)3-5/h1-3,8,10-12H,4,9H2/t8-/m0/s1')
```

You can also use an InChI key:
```{r}
chebi$convInchiToChebi('MBDOYVRWFFCFHM-SNAWJCMRSA-N')
```

# Closing biodb instance

Do not forget to terminate your biodb instance once you are done with it:
```{r}
mybiodb$terminate()
```
